### 2025-01-28 scripts/Paper-analysis.Rmd
sliding window figures

# Maximum counts
d.reduced.median.plot = d.reduced.median %>% 
  group_by(PTU) %>%
  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
  ungroup() %>%  # Remove the grouping
  select(-max_n)  # Optionally remove the max_n column if not needed
# Plot the distribution
p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median))+
  geom_hline(yintercept = 0, linetype='dashed')+
  theme_bw()+
  geom_path(aes(group=PTU, alpha=normalized_n))+
  facet_wrap(~PTU, scales="free_x", nrow=3)+
  ylab("Deviation from expected 6-mer palindrome density (5kb sliding window)")+
  labs(alpha="% of plasmids within PTU")+
  theme(panel.grid=element_blank())+  
  scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(legend.position = c(0.8, 0.2))
ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf',
       width=10, height=6)
```

Repeat for k=4?

>>>>

# Maximum counts
d.reduced.median.plot = d.reduced.median %>% 
  group_by(PTU) %>%
  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
  ungroup() %>%  # Remove the grouping
  select(-max_n)  # Optionally remove the max_n column if not needed
# Plot the distribution
p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median*1000))+
  geom_hline(yintercept = 0, linetype='dashed')+
  theme_bw()+
  geom_path(aes(group=PTU, alpha=normalized_n))+
  facet_wrap(~PTU, scales="free_x", nrow=3)+
  ylab("Palindrome depletion per kb (5kb sliding window)")+
  labs(alpha="% of plasmids within PTU")+
  theme(panel.grid=element_blank())+  
  scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(legend.position = c(0.8, 0.2))
ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf',
       width=10, height=6)
# Plot only first 25kb
ggsave(p.lines.expectation+scale_x_continuous(labels = function(x) x / 1000, limits=c(0,50000)), file='../results/6mer-deviation-expectation-0-50-kb.pdf',
       width=10, height=6)

### Compare leading region to rest of plasmid

```{r}
# Start from d.reduced, which contains per-plasmid results with 5000bp sliding window and step size of 500bp
# Compare depletion in first 5kb with average for the rest of the plasmid
diff.leading.region.average = d.reduced %>% 
  group_by(plasmid, PTU) %>%
  mutate(diff.count=palindrome_count[x=="0"]-mean(palindrome_count),
         diff.density=diff_density[x=="0"]-mean(diff_density)) %>%
  filter(x=="0")

# Order the PTUs by mean values of depletion for better visual display
mean.values = diff.leading.region.average %>% group_by(PTU) %>%
  summarise(mean=mean(diff.density)*1000)
diff.leading.region.average$PTU = ordered(diff.leading.region.average$PTU,
                                          levels=mean.values$PTU[order(mean.values$mean)])
p.leading.region.diff.density = ggplot(diff.leading.region.average, aes(as.factor(PTU), diff.density*1000))+ # depletion per-kb rather than per-bp
  ggbeeswarm::geom_quasirandom(alpha=0.8)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_hline(yintercept = 0)+
  xlab("")+
  theme(axis.text=element_text(colour="black"))+
  ylab("Palindrome depletion in leading region (per kb)")+
  coord_flip()
p.leading.region.diff.density
# Based on this, remove two outliers
p.leading.region.diff.density = p.leading.region.diff.density+
  ylim(c(-10,5))
ggsave(p.leading.region.diff.density, file='../results/6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
       width=5, height=4)
```
